\documentclass[11pt,letterpaper]{article}

% ── Geometry ──────────────────────────────────────────────────────────────────
\usepackage[margin=1in]{geometry}

% ── Pagination ────────────────────────────────────────────────────────────────
\usepackage[all]{nowidow}    % Require ≥2 lines on each side of a page break
\brokenpenalty=10000         % No hyphenated words across page breaks
\raggedbottom                % Allow uneven page bottoms rather than stretching

% ── Fonts ─────────────────────────────────────────────────────────────────────
\usepackage[T1]{fontenc}
\usepackage{newpxtext}     % Palatino-based serif (TeX Gyre Pagella)
\usepackage{newpxmath}     % Matching math font

% ── Colors ────────────────────────────────────────────────────────────────────
\usepackage[dvipsnames]{xcolor}
\definecolor{SectionBlue}{HTML}{1B3A5C}
\definecolor{AccentGray}{HTML}{4A4A4A}
\definecolor{QuoteBorder}{HTML}{8B9DAF}
\definecolor{RiskRed}{HTML}{C0392B}
\definecolor{RiskAmber}{HTML}{E67E22}
\definecolor{RiskGreen}{HTML}{27AE60}
\definecolor{BoxBg}{HTML}{F7F8FA}

% ── Tables ────────────────────────────────────────────────────────────────────
\usepackage{booktabs}
\usepackage{tabularx}

% ── Boxes ─────────────────────────────────────────────────────────────────────
\usepackage[framemethod=tikz]{mdframed}

% ── Lists ─────────────────────────────────────────────────────────────────────
\usepackage{enumitem}
\usepackage{needspace}

% ── Hyperlinks ────────────────────────────────────────────────────────────────
\usepackage{hyperref}
\hypersetup{
  colorlinks=true,
  linkcolor=SectionBlue,
  urlcolor=SectionBlue,
  citecolor=SectionBlue,
  pdfauthor={Pat Singh},
  pdftitle={PR/FAQ — Beadle},
  pdfsubject={Working Backwards — Product Discovery},
}

% ── Bibliography ─────────────────────────────────────────────────────────────
\usepackage[style=numeric,backend=biber,sorting=none]{biblatex}
\addbibresource{prfaq.bib}

% ── Section styling ───────────────────────────────────────────────────────────
\usepackage{titlesec}
\titleformat{\section}
  {\needspace{6\baselineskip}\Large\bfseries\color{SectionBlue}}
  {}
  {0pt}
  {}

\titleformat{\subsection}
  {\needspace{5\baselineskip}\large\bfseries\color{AccentGray}}
  {}
  {0pt}
  {}

\titleformat{\subsubsection}
  {\normalsize\bfseries\color{AccentGray}}
  {}
  {0pt}
  {}
\titlespacing*{\subsubsection}{0pt}{1em}{0.5em}[0pt]

% ── Document stage ───────────────────────────────────────────────────────────
\newcommand{\prfaqstage}[1]{\def\theprfaqstage{#1}}
\prfaqstage{hypothesis}

% ── Document version ─────────────────────────────────────────────────────────
\newcommand{\prfaqversion}[2]{\def\theprfaqmajor{#1}\def\theprfaqminor{#2}}
\prfaqversion{3}{0}

% ── Header/footer ────────────────────────────────────────────────────────────
\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{}
\fancyhead[L]{\footnotesize\textcolor{AccentGray}{Stage: \theprfaqstage{} \enspace\textbar\enspace v\theprfaqmajor.\theprfaqminor}}
\fancyhead[R]{\footnotesize\textcolor{AccentGray}{\thepage}}
\renewcommand{\headrulewidth}{0pt}

% ══════════════════════════════════════════════════════════════════════════════
% Custom environments
% ══════════════════════════════════════════════════════════════════════════════

% ── \prsection{title} — styled press release section header ──────────────────
\newcommand{\prsection}[1]{%
  \vspace{1em}%
  {\large\bfseries\color{SectionBlue}#1}%
  \vspace{0.3em}\par%
}

% ── customerquote — indented, italic customer quote ──────────────────────────
\newmdenv[
  topline=false,
  bottomline=false,
  rightline=false,
  linewidth=3pt,
  linecolor=QuoteBorder,
  backgroundcolor=BoxBg,
  innerleftmargin=12pt,
  innerrightmargin=12pt,
  innertopmargin=8pt,
  innerbottommargin=8pt,
  skipabove=\baselineskip,
  skipbelow=\baselineskip,
]{customerquote}

% ── spokespersonquote — indented spokesperson quote ──────────────────────────
\newmdenv[
  topline=false,
  bottomline=false,
  rightline=false,
  linewidth=3pt,
  linecolor=SectionBlue,
  backgroundcolor=BoxBg,
  innerleftmargin=12pt,
  innerrightmargin=12pt,
  innertopmargin=8pt,
  innerbottommargin=8pt,
  skipabove=\baselineskip,
  skipbelow=\baselineskip,
]{spokespersonquote}

% ── faqpair — numbered Q&A pair with bold question ─────────────────────────
\newcounter{faqnum}
\newenvironment{faqpair}[1]{%
  \needspace{4\baselineskip}%
  \vspace{0.5em}%
  \refstepcounter{faqnum}%
  \noindent\textbf{\color{SectionBlue}Q\thefaqnum: #1}\par\nopagebreak[4]%
  \vspace{0.2em}%
  \begin{adjustwidth}{1em}{0pt}%
  \setlength{\parindent}{0pt}%
}{%
  \end{adjustwidth}%
  \vspace{0.5em}%
}
% ── \faqref{faq:slug} — clickable "FAQ 3" reference ──────────────────────────
\newcommand{\faqref}[1]{\hyperref[#1]{FAQ~\ref*{#1}}}

% ── \featureitem — numbered, referenceable feature entry ──────────────────────
\newcounter{featurenum}
\newcommand{\featureitem}[2]{%
  \refstepcounter{featurenum}%
  \item[\textbf{\color{SectionBlue}F\thefeaturenum.}]%
  \textbf{#1} --- #2%
}
\newcommand{\featureref}[1]{\hyperref[#1]{Feature~\ref*{#1}}}

% ── adjustwidth (for faqpair indentation) ────────────────────────────────────
\usepackage{changepage}

% ══════════════════════════════════════════════════════════════════════════════
% Document
% ══════════════════════════════════════════════════════════════════════════════

\begin{document}

% ── Headline block ───────────────────────────────────────────────────────────
\begin{center}
  {\LARGE\bfseries\color{SectionBlue}
    Punt Labs Launches Beadle} \\[0.8em]
  {\large\color{AccentGray}
    Signed instructions. Declared permissions. Tamperproof audit trail.\\
    Your autonomous agent, on your machine, under your control.}
\end{center}

\vspace{0.5em}

% ══════════════════════════════════════════════════════════════════════════════
% PRESS RELEASE
% ══════════════════════════════════════════════════════════════════════════════

\section*{Press Release}

SEATTLE, WA --- September 15, 2026 --- Punt Labs today released Beadle,
a free, open-source daemon that turns email into a programmable control
plane for an autonomous agent running on your own machine. You send an
email; Beadle verifies your identity via PGP or Proton E2E encryption,
decomposes your instruction into a pipeline of commands, and executes
them --- mixing AI reasoning (Claude) with fast CLI tools (biff, jq,
git) in a single pipeline. Results arrive as an email reply. Unlike
agent frameworks that ask for broad trust, Beadle requires GPG-signed
command definitions, enforces per-contact permissions before executing
anything, and records every step in a pipeline audit trail. Commands
are programs, the daemon is the shell, pipelines are pipes, and GPG
signatures are sudo (see \faqref{faq:competitors}).

\prsection{Problem}

Developers today use AI coding assistants interactively: they sit at the
keyboard, type a prompt, review the output, and iterate. When the laptop closes,
the agent stops. A developer who triggers a two-hour test suite at 6pm must
either keep the terminal open until 8pm or accept that a crash at 7pm means
starting over in the morning. Delegating multi-step tasks --- run tests across
three environments, summarize the results, file issues for failures --- means
cobbling together fragile cron jobs and shell scripts with no authentication, no
permissions model, and no audit trail. Existing autonomous agent frameworks like Auto-GPT run locally but offer
no cryptographic verification of who authorized which action, no per-command
permissions, and no tamperproof record of what happened.
The risk is not theoretical: in January 2026, ClawdBot (OpenClaw) gained
85,000+ GitHub stars in its first week and was immediately found to have
critical vulnerabilities --- authentication disabled by default, a skill
directory riddled with malicious extensions, and prompt injection paths that
exfiltrated API keys \cite{kaspersky2026_openclaw, snyk2026_clawdbot}.
Autonomous agents without enforceable trust boundaries produce real security
failures at scale (see \faqref{faq:customer-evidence}).

\prsection{Solution}

Beadle is a background daemon with its own email address, GPG keychain,
and a library of GPG-signed YAML command files. Each command declares
a runner (Claude for AI reasoning, CLI for deterministic tools), its
inputs and outputs via JSON schema, and whether it transforms the pipeline
data or passes it through as a side-effect. Commands compose like Unix
pipes:

{\small\texttt{summarize | notify-team | reply}}

The owner sends an email to Beadle's address. The daemon verifies the
sender's identity (PGP signature or Proton E2E encryption), checks their
\texttt{x} (execute) permission, maps the instruction to a command
pipeline, and executes stages sequentially. A single JSON value --- the
pipe --- flows through the pipeline. Process-mode commands transform it;
passthrough-mode commands (notifications, logging) read it without
modification. Results arrive as an email reply. Every stage is logged
with pipeline ID, command name, runner, duration, and exit status
(see \faqref{faq:why-email}).

\prsection{Customer Quote}

\begin{customerquote}
\textit{``I emailed `summarize this' to my agent and 90 seconds later had
a structured summary in my inbox. The Claude stage did the reasoning, a
jq + biff stage broadcasted the headline to my team in 10 milliseconds,
and the reply landed in my inbox with the full analysis. Three stages,
two different runners, one email. I didn't have to be at my keyboard.
The pipeline log shows exactly what happened: which command, which runner,
how long, what output. It's like having cron + Unix pipes + GPG sudo for
an AI agent.''}

\raggedleft --- \textbf{Alex Petrov}, Senior Platform Engineer
\end{customerquote}

\prsection{Getting Started}

Getting started takes three steps. First, install the binary:
{\small\texttt{curl -fsSL .../install.sh | sh}} downloads the Go binary
and installs the Claude Code plugin (MCP tools + slash commands). Second,
configure your email connection (IMAP/SMTP via Proton Bridge or any
provider) and store credentials in your OS keychain. Third, start the
daemon: {\small\texttt{beadle-daemon run}}. Beadle polls your inbox,
and when you send an email with ``summarize'' in the subject, you get
a structured summary back within two minutes. No cloud account, no data
leaves your machine. Commands are YAML files in
{\small\texttt{\textasciitilde/.punt-labs/beadle/commands/}} --- add new
capabilities by writing a YAML file and signing it with your GPG key.

\prsection{Spokesperson Quote}

\begin{spokespersonquote}
``At Punt Labs we build on a principle: earn trust to go fast.
Every autonomous agent framework asks you to trust it implicitly --- give it
your credentials, let it run whatever it wants, and hope the logs are honest.
We built Beadle on the opposite assumption: the agent has zero authority
of its own. Every signed instruction, every declared permission, every
tamperproof log entry earns the right to automate more aggressively tomorrow.
The signing ceremony isn't overhead --- it's the mechanism that compounds trust
so you can hand off bigger tasks with confidence. It's the Unix philosophy
applied to AI agents: small, composable, auditable tools that do one thing
well.''

\raggedleft --- \textbf{Pat Singh}, Punt Labs
\end{spokespersonquote}

\prsection{Call to Action}

Beadle is available now at
{\small\texttt{github.com/punt-labs/beadle}} under the MIT license.
Static Go binaries for macOS and Linux (arm64/amd64). Installation
requires an IMAP/SMTP server (Proton Bridge recommended for E2E
encryption) and GPG for command signing. Documentation, example
commands, and the pipeline design document are in the repository.

\newpage

% ══════════════════════════════════════════════════════════════════════════════
% FREQUENTLY ASKED QUESTIONS
% ══════════════════════════════════════════════════════════════════════════════

\section*{Frequently Asked Questions}

% ── External FAQs (Customer-facing) ──────────────────────────────────────────

\subsection*{External FAQs}

\begin{faqpair}{What is Beadle and who is it for?}\label{faq:what-is-it}
  Beadle is an open-source agent daemon that receives instructions via email
  and executes them as multi-stage pipelines. It is for developers who want
  to delegate tasks to an AI agent running on their own machine --- with
  cryptographic verification of who authorized what. Commands are YAML files
  that declare a runner (Claude for reasoning, CLI for tools like jq, biff,
  git), a JSON output schema, and whether they transform the pipeline data
  or pass it through as a side-effect. The daemon is the shell; commands
  are programs; GPG signatures are sudo.
\end{faqpair}

\begin{faqpair}{How is Beadle different from Auto-GPT or Claude Code?}\label{faq:competitors}
  Auto-GPT and similar frameworks give an agent broad autonomy with minimal
  authentication. Claude Code is an interactive assistant that requires the
  developer at the keyboard (its \texttt{-{}-bare} mode enables headless
  execution but provides no trust model, no command library, and no pipeline
  orchestration).

  Beadle occupies a different point: it has \textit{zero} independent authority.
  Every command is a GPG-signed YAML file with a declared runner, typed args,
  and JSON output schema. Pipelines compose explicitly. The trust gate verifies
  sender identity (PGP or Proton E2E) and per-contact permissions before any
  execution. CLI commands run in milliseconds with whitelist-only binary
  resolution and no shell. The trade-off is deliberate: Beadle sacrifices agent
  autonomy for owner control, auditability, and speed on deterministic operations.
\end{faqpair}

\begin{faqpair}{Why email? Why not a web UI or a CLI?}\label{faq:why-email}
  Email is asynchronous, universal, and already authenticated. Proton Mail
  provides end-to-end encryption. GPG provides sender verification. Together,
  they give Beadle a control plane that works from any device --- phone, tablet,
  another computer --- without installing a client or opening a port. The
  concept of email as an agent control channel has engineering precedent:
  HumanLayer's Agent Control Plane uses email for human-approval of autonomous
  agent actions \cite{humanlayer2024acp}. The owner
  can CC Beadle on a conversation with a colleague, and Beadle can distinguish
  the owner's signed instructions from the colleague's unsigned replies.

  Beadle also supports local execution via the CLI
  ({\small\texttt{beadle run pipeline.md}}) for development and testing.
  Email is the remote interface, not the only interface.
\end{faqpair}

\begin{faqpair}{How do I get started?}\label{faq:getting-started}
  Run the install script ({\small\texttt{curl -fsSL .../install.sh | sh}})
  to download the Go binary and install the Claude Code plugin. Configure
  your email connection (IMAP/SMTP) and credentials. Start the daemon with
  {\small\texttt{beadle-daemon run}}. Send an email with ``summarize'' in
  the subject from a PGP-verified or Proton E2E contact, and receive the
  summary as a reply. Add commands by writing YAML files in
  {\small\texttt{\textasciitilde/.punt-labs/beadle/commands/}}.
\end{faqpair}

\begin{faqpair}{How much does it cost?}\label{faq:pricing}
  Beadle is free and open-source. You need a Proton Mail paid plan
  (starting at \$3.99/month) because Proton Bridge --- the local IMAP/SMTP
  server Beadle uses for email --- is not available on the free tier
  \cite{protonbridge2025}. You also need a GPG keychain (free). If your command
  pipelines use the Claude interpreter, you will need an Anthropic API key with
  usage-based billing. Beadle itself has no cost.
\end{faqpair}

\begin{faqpair}{What happens to my data?}\label{faq:data}
  Nothing leaves your machine unless a command document explicitly sends it
  somewhere. Beadle runs locally as a daemon. Email flows through Proton Bridge
  on your machine to Proton's encrypted servers. Command documents, audit logs,
  and results stay on your local filesystem. There is no telemetry, no cloud
  service, and no account with Punt Labs.
\end{faqpair}

\begin{faqpair}{Do I need to know GPG?}\label{faq:gpg-knowledge}
  No. {\small\texttt{beadle init}} generates a signing key in an isolated
  keychain (separate from your system GPG keyring), configures Proton Bridge,
  and produces a single identity file. You never run \texttt{gpg} directly ---
  {\small\texttt{beadle sign}} and {\small\texttt{beadle verify}} handle all
  cryptographic operations. If you already have a GPG key you want to use,
  Beadle can import it, but it still stores it in its own keychain. The only
  requirement is that signing keys must have an expiration date ---
  Beadle enforces this and will not accept non-expiring keys for command
  signing.
\end{faqpair}

% ── Internal FAQs (Business-facing) ──────────────────────────────────────────

\newpage
\subsection*{Internal FAQs}

\subsubsection*{Value \& Market}

\begin{faqpair}{What is the total addressable market?}\label{faq:tam}
  The primary market is developers who already use AI coding assistants and want
  unattended execution. The LangChain 2024 State of AI Agents report surveyed
  1,300+ professionals and found that 51\% are using AI agents in production
  today, with 78\% planning to deploy agents soon \cite{langchain2024agents}.
  The subset who need local-first, authenticated, autonomous execution is
  narrower --- likely tens of thousands of developers who work with sensitive
  codebases, regulated environments, or privacy-conscious infrastructure
  (the Proton Mail user base is an imperfect proxy for this audience).

  This is a bottoms-up estimate. We are not claiming a billion-dollar TAM.
  The honest framing: Beadle targets a niche of security-conscious developers
  who value auditability over convenience.
\end{faqpair}

\begin{faqpair}{What evidence do we have that customers want this?}\label{faq:customer-evidence}
  At hypothesis stage, we have limited primary data but one working prototype.
  The evidence:

  \textbf{Primary evidence (owner use):} A proof-of-concept runs daily on the
  author's machine. A GPG-signed prompt policy manifest
  ({\small\texttt{prompts.policy.yml}}) defines a healthcheck command with
  SHA-256 integrity verification, trusted key enforcement, per-command
  constraints (model, max cost, allowed tools, I/O boundaries), and template
  variable substitution. A shell wrapper verifies the GPG signature, invokes
  Claude non-interactively, and delivers results by email via Proton Bridge.
  The agent has its own Proton Mail address and delivers results via Proton
  Bridge. This prototype has run reliably as a daily LaunchAgent for several
  weeks, validating the core integration: GPG-signed prompts, Claude CLI
  invocation, Proton Bridge email delivery, and scheduled execution.

  \textbf{Inferential evidence:}
  \begin{itemize}[nosep]
    \item Auto-GPT has accumulated over 182,000 GitHub stars since its 2023
      release, demonstrating strong developer interest in autonomous agents
      \cite{autogpt2024github}.
    \item The OWASP GenAI Security Project published a Top~10 for Agentic
      Applications in 2025, indicating that agent security is a recognized
      concern, not a fringe worry \cite{owasp2025agentic}.
    \item Proton Mail has grown to 100M+ accounts, confirming demand for
      privacy-focused communication infrastructure \cite{proton2024100m}.
    \item Cisco's 2025 Data Privacy Benchmark found that 90\% of
      organizations see local storage as inherently safer, and 64\% worry
      about exposing sensitive data to cloud GenAI tools
      \cite{cisco2025privacy}.
    \item No existing agent framework offers GPG-based authentication with
      per-command permissions --- the gap is observable, but unvalidated.
  \end{itemize}

  \textbf{Next validation step:} Generalize the prototype into the Beadle
  daemon (multi-command pipelines, identity management via
  {\small\texttt{beadle init}}). Then test with 5--10 external developers
  who currently use Auto-GPT or Claude Code for automation tasks. Measure
  whether the GPG authentication model feels like a feature or a burden.
\end{faqpair}

\begin{faqpair}{Who are the competitors and why will we win?}\label{faq:competitive-landscape}
  \begin{tabularx}{\linewidth}{@{} l X X @{}}
  \toprule
  \textbf{Competitor} & \textbf{Strength} & \textbf{Trust model gap} \\
  \midrule
  Auto-GPT & Large community, broad agent autonomy &
    Agent-directed execution; no owner authentication, no declared
    permissions, no tamperproof audit trail \\
  Claude Code & High-quality code generation, interactive &
    Operator-attended trust model; requires user at keyboard,
    cannot delegate unattended authority \\
  Devin & Full autonomous dev environment &
    Vendor-hosted trust model; execution runs on Cognition's
    infrastructure, not the owner's machine \\
  Cron + scripts & Simple, well-understood &
    No trust model; no identity verification, no permission
    scoping, no composability with LLMs \\
  \bottomrule
  \end{tabularx}

  \vspace{0.5em}
  Beadle does not compete on agent intelligence --- it uses existing LLMs
  (Claude, GPT) as interpreters. The differentiation is the trust model:
  local-only execution under cryptographic owner control. Cloud providers and
  SaaS agent platforms cannot replicate this without contradicting their own
  business model --- their revenue depends on hosting execution, managing
  infrastructure, and retaining custody of the runtime environment. Offering
  genuine local-only, owner-controlled execution means giving up the recurring
  billing, telemetry, and platform lock-in that sustain their economics.
  Competitors can add daemon modes, unattended execution, and even signing
  features, but they cannot cede control of the execution environment to the
  owner without dismantling the value capture that funds their products. This
  is a structural misalignment, not a feature gap.
\end{faqpair}

\begin{faqpair}{What is the customer acquisition strategy?}
  Open-source distribution via GitHub Releases. Beadle is a CLI tool in the Punt
  Labs ecosystem (alongside Biff, Quarry, and Vox), which provides cross-project
  visibility. The secondary purpose --- as a showcase for the Jacobson-Cockburn
  use-case methodology \cite{jacobson2023usecases} --- generates content that
  attracts developers interested in both the tool and the process.

  No paid acquisition. Growth is organic: one launch post (Hacker News,
  Lobsters), 2--3 blog posts documenting the build process and methodology,
  and cross-links from other Punt Labs project READMEs. Estimated effort:
  20--30 hours of content creation in the first 3 months. If that generates
  fewer than 100 stars, the positioning missed the audience and content
  investment should stop.
\end{faqpair}

\begin{faqpair}{What is your next step to validate your vision?}\label{faq:next-step}
  The daemon is running. Pipeline v1 is verified end-to-end: a real email
  from a PGP/Proton-verified sender triggers a multi-stage pipeline
  (summarize via Claude, auto-reply via Claude), and the sender receives
  the result. 9 PRs merged, 3 security reviews, e2e tested on Proton Bridge.

  The next step is Pipeline v2: add CLI runners (execute tools like biff,
  jq, git in milliseconds instead of 45-second Claude sessions),
  process/passthrough data flow (side-effect commands don't destroy the
  pipe), compound CLI steps (jq | biff chained without a shell), and
  JSON schema validation between stages. The design has passed three
  rounds of architect review and is broken into 8 implementation tasks.

  After v2 ships, the tool is ready for external testing: 5--10 developers
  who use Claude Code or Auto-GPT for automation. The test: can they
  write a command YAML, sign it, send an email, and receive a pipeline
  result without consulting GPG documentation? The key question is
  whether the trust model feels like a feature or a burden.
\end{faqpair}

\subsubsection*{Technical}

\begin{faqpair}{What are the major technical risks?}\label{faq:tech-risks}
  \begin{enumerate}[nosep]
    \item \textbf{Proton Bridge reliability for automated use.} Proton Bridge
      is designed for desktop email clients, not daemons. It may not handle
      long-running, unattended operation gracefully.

      \textit{Why Proton specifically:} Beadle's threat model assumes the
      email channel must protect command integrity end-to-end --- from the
      owner's device to the daemon's mailbox. Proton Mail's end-to-end
      encryption means even Proton's own servers cannot read command content in
      transit, which aligns with the local-first, zero-trust design. That said,
      Beadle connects to Proton Bridge via standard IMAP/SMTP --- there is no
      Proton-specific API or protocol dependency. Supporting other email
      providers (Fastmail, self-hosted Dovecot, Gmail with app passwords) is a
      scope decision, not an architectural constraint.

      \textit{Mitigation:} Beadle monitors Bridge health on a configurable
      interval (default: 60 seconds) and exposes status via
      {\small\texttt{beadle status}}. If Bridge becomes unavailable, Beadle
      logs the outage, queues pending result deliveries, and retries on a
      backoff schedule. On macOS, the LaunchAgent plist includes a
      {\small\texttt{KeepAlive}} directive that restarts the daemon
      automatically if it exits; on Linux, the systemd unit specifies
      {\small\texttt{Restart=on-failure}} with a 30-second delay. The owner
      can configure alerting (e.g., a local notification or a write to a
      monitored file) when Bridge has been offline for longer than a threshold.
      Email-triggered execution is unavailable while Bridge is down, but
      locally triggered and cron-triggered pipelines continue to run and their
      results are delivered once Bridge recovers.
    \item \textbf{GPG key management complexity.} GPG tooling is notoriously
      difficult --- complexity is consistently cited as its primary operational
      weakness \cite{hoopdotdev2024gpg}. If {\small\texttt{beadle init}} cannot
      fully abstract key generation and signing, the onboarding friction will
      kill adoption. This is the hardest unsolved UX problem in the project.

      \textit{Alternative evaluated: SSH signing.} GitHub and GitLab both
      support SSH keys for commit signing, and most developers already have an
      SSH key. However, SSH signing only covers commit-level attestation ---
      it provides no standard for encrypting documents to a recipient, no
      built-in key expiry or revocation model, and no ecosystem for
      distributing trust (no equivalent of the OpenPGP Web of Trust or
      keyservers). Beadle needs signing \textit{and} encryption \textit{and}
      key lifecycle management; SSH covers only the first.

      \textit{Mitigation:} {\small\texttt{beadle init}} must make GPG
      invisible. Concretely, it must: (1)~generate an Ed25519 signing key
      with a mandatory expiration date, storing it in an isolated
      {\small\texttt{GNUPGHOME}} so it does not pollute the user's existing
      keychain; (2)~configure Proton Bridge IMAP/SMTP credentials and verify
      connectivity with a round-trip test message; (3)~produce a single
      {\small\texttt{beadle.yml}} identity file that all subsequent commands
      consume. The user should never run \texttt{gpg} directly.
      {\small\texttt{beadle sign}} wraps signing, and
      {\small\texttt{beadle verify}} wraps verification. If this abstraction
      leaks --- if any error message surfaces raw \texttt{gpg} output or
      requires the user to understand keyring semantics --- the onboarding
      has failed.
    \item \textbf{Claude interpreter sandboxing.} When Beadle invokes Claude
      as a command interpreter, the LLM may attempt actions outside the
      command's declared permissions. \textit{Mitigation:} The two-level
      try-catch model preflights all permissions before execution. The Claude
      interpreter runs in a restricted context with only the declared
      inputs available.
  \end{enumerate}
\end{faqpair}

\begin{faqpair}{What dependencies exist on other teams or systems?}
  \begin{itemize}[nosep]
    \item \textbf{Proton Bridge} --- open-source Go codebase, maintained by
      Proton AG \cite{protonbridge2025}. Beadle uses it as a standard
      IMAP/SMTP client. No API dependency beyond the mail protocol.
    \item \textbf{Anthropic API} --- required only when using the Claude
      interpreter. Usage-based billing, no contractual dependency.
    \item \textbf{GnuPG} --- the GPG implementation. Mature, widely deployed,
      maintained by the GnuPG Project. RFC~9580 (OpenPGP, 2024) provides
      the current standard \cite{rfc9580}.
  \end{itemize}
  No internal Punt Labs dependencies. Beadle is a standalone project.
\end{faqpair}

\begin{faqpair}{What is the estimated development timeline?}\label{faq:timeline}
  This is a hypothesis-stage document. The purpose of Working Backwards is
  to validate \textit{whether} to build, not to plan the schedule. Detailed
  hour estimates and calendar targets are premature --- and doubly so when
  AI-augmented development makes historical velocity assumptions unreliable.
  What matters at this stage is delivery order and cut discipline.

  \textbf{Phased delivery order:}

  \begin{enumerate}[nosep]
    \item \textbf{Foundation (shipped)} --- MCP email server with 17 tools,
      four-level PGP trust model, per-contact rwx permissions, multi-identity
      via ethos, credential isolation (OS keychain), Proton Bridge IMAP/SMTP.
      Released as \texttt{beadle-email} (v0.14.1).
    \item \textbf{Daemon + Pipeline v1 (shipped)} --- Background daemon with
      IMAP poller, trust gate (PGP verified + Proton E2E), x-bit enforcement,
      pipeline orchestrator (command YAML loader, regex planner, sequential
      executor, auto-reply, crash-safe persistence), worker spawner via
      \texttt{claude -p --bare} with adversarial isolation. 9 PRs, 3 security
      reviews, e2e verified on Proton Bridge.
    \item \textbf{Pipeline v2 (in progress)} --- CLI runners (execute tools
      in milliseconds), process/passthrough data flow, compound CLI steps
      (jq | biff without shell), JSON pipe with schema validation,
      runner-conditional command validation. 8 tasks across 4 phases.
      Three architect reviews completed.
    \item \textbf{Future} --- LLM planner (replace regex with Claude
      decomposition), GPG command signing enforcement, multi-channel triggers
      (biff chat, webhooks), internal scheduler, key rotation management.
  \end{enumerate}

  \textbf{Cut order:} Future items are cut first. Pipeline v2 is the
  current priority --- it enables the command library that makes beadle
  useful beyond the ``summarize'' demo. The kill switch is adoption-based:
  see \faqref{faq:metrics}.
\end{faqpair}

\begin{faqpair}{What is the scaling story?}\label{faq:scaling}
  Beadle is a single-user daemon. ``Scaling'' means long-running operation over
  months and years, not multi-tenant growth.

  \begin{itemize}[nosep]
    \item \textbf{Audit log growth.} The log is append-only and unbounded until
      the owner manually clears it. A developer running 10 pipelines per day
      generates roughly 50--100~KB of log per day. At one year, that is
      approximately 20--40~MB --- manageable on any modern filesystem. The risk
      is neglect, not size.
    \item \textbf{Command library size.} Commands are loaded into memory only
      during preflight and execution, then discarded. A library of hundreds of
      signed command documents on disk has no runtime cost until invoked.
    \item \textbf{Proton Bridge longevity.} The daemon depends on Bridge staying
      online for email-triggered execution. Bridge was not designed for months
      of continuous uptime. The mitigation is operational: Beadle checks Bridge
      health on a configurable interval, the system service manager (LaunchAgent
      or systemd) restarts both Beadle and Bridge on failure, and pending result
      deliveries are queued and retried once Bridge recovers. Local and
      cron-triggered pipelines are unaffected by Bridge downtime.
  \end{itemize}

  The honest constraint: Beadle does not scale to teams, organizations, or
  multi-machine deployment. It is a personal agent for one person on one
  machine. That is a design choice, not a limitation to fix later.
\end{faqpair}

\subsubsection*{Business}

\begin{faqpair}{What is the revenue model?}\label{faq:revenue}
  None. Beadle is free and open-source. The strategic value is threefold:
  (1)~it demonstrates the Punt Labs engineering methodology in a real,
  useful product; (2)~it grows the Punt Labs open-source ecosystem alongside
  Biff, Quarry, and TTS; (3)~it serves as the reference implementation for
  a future punt-labs project on use-case methodology, generating content and
  credibility.
\end{faqpair}

\begin{faqpair}{What does the P\&L look like at steady state?}\label{faq:pnl}
  Beadle has no revenue, so the P\&L is cost-only. The primary cost is
  developer-hours. Specific hour estimates are not attempted at hypothesis
  stage --- the purpose of this document is to validate whether to build,
  not to plan the schedule --- but the cost categories are known.

  \begin{tabularx}{\linewidth}{@{} l X @{}}
  \toprule
  \textbf{Cost} & \textbf{Notes} \\
  \midrule
  Development (primary) & Solo developer. Four phases described in
    \faqref{faq:timeline}. The dominant cost and the binding constraint
    on opportunity cost. \\
  Ongoing maintenance & Bug fixes, dependency updates, community PRs.
    Modest but nonzero indefinitely. \\
  Proton Mail (paid plan) & \$48/year. Personal use; already paid for
    other purposes. \\
  Infrastructure & \$0. No servers; runs on developer's machine. \\
  \bottomrule
  \end{tabularx}

  \vspace{0.5em}
  The opportunity cost is the binding constraint: development time spent on
  Beadle is time not spent on other Punt Labs projects (Quarry features, Biff
  improvements, LangLearn ports). This cost is governed by a kill switch tied
  to the metrics in \faqref{faq:metrics}: if, at the 6-month mark after
  public release, Beadle has fewer than 100 GitHub stars \textit{or} fewer
  than 10 weekly binary downloads, the project enters archival mode (security
  patches only, no new feature work). This is a hard threshold, not a
  judgment call --- below either number means the positioning missed the
  audience and further development hours are not justified.
\end{faqpair}

\begin{faqpair}{What are the key metrics and how will we measure success?}\label{faq:metrics}
  \begin{tabularx}{\linewidth}{@{} l l X @{}}
  \toprule
  \textbf{Metric} & \textbf{Target (6 months)} & \textbf{Decision threshold} \\
  \midrule
  GitHub stars & 500+ & Below 100 after launch suggests the positioning
    missed the audience \\
  Weekly binary downloads & 50+ & Below 10 suggests the tool is not
    useful enough to keep \\
  Prototype test completion & 8/10 users complete first pipeline &
    Below 5/10 signals fatal usability friction \\
  \bottomrule
  \end{tabularx}

  \vspace{0.5em}
  \textbf{Pre-mortem --- what failure looks like:} It is March 2027. Beadle has
  200 GitHub stars, 8 binary downloads per week, and the prototype test showed
  3/10 users abandoned during GPG key setup. The cryptographic trust model ---
  signed commands, declared permissions, tamperproof audit trail --- resonated
  in blog posts but not in practice. Developers found the signing ceremony
  too heavy for personal automation tasks where they already trust themselves.
  The security posture Beadle enforces felt like overhead, not protection.
  The project becomes a methodology showcase only, with no active user
  community, and enters archival maintenance.
\end{faqpair}

\begin{faqpair}{Why now? What has changed?}\label{faq:why-now}
  Four shifts make this the right time:

  \begin{enumerate}[nosep]
    \item \textbf{AI agents went mainstream in 2024--2025.} The LangChain survey
      shows 51\% production deployment \cite{langchain2024agents}. Developers
      are past the question of ``should I use an agent?'' and into ``how do I
      trust one?''
    \item \textbf{OpenPGP modernized.} RFC~9580 (2024) updated the OpenPGP
      standard for the first time in decades, adding modern cryptographic
      algorithms and improving key management \cite{rfc9580}. GPG tooling is
      better than it was even two years ago.
    \item \textbf{OWASP formalized agent security risks.} The OWASP Top~10 for
      Agentic Applications (2025) legitimized the concern that autonomous agents
      need security models, not just capability models
      \cite{owasp2025agentic}. Beadle's trust model is a direct response.
    \item \textbf{ClawdBot proved the threat model in production.} In January
      2026, ClawdBot (also known as OpenClaw) launched and gained 85,000+
      GitHub stars in its first week. Within days, security researchers
      discovered critical vulnerabilities: CVE-2026-25253 (CVSS~8.8) allowed
      gateway compromise via authentication bypass in misconfigured reverse
      proxies, CVE-2026-24763 enabled hundreds of malicious skills in the skill
      directory, and CVE-2026-25157 allowed prompt injection to extract secrets
      \cite{kaspersky2026_openclaw}. Snyk demonstrated that a spoofed email
      could trigger a shell command that exfiltrated \texttt{clawdbot.json}
      containing API keys \cite{snyk2026_clawdbot}. Palo Alto Networks'
      Unit~42 concluded that OpenClaw ``maps to all 10 OWASP risks for agentic
      applications'' and lacked ``enforceable trust boundaries between untrusted
      inputs and high-privilege reasoning'' \cite{unit42_2026_openclaw}.
      Kaspersky reported that authentication was disabled by default
      \cite{kaspersky2026_openclaw}. This is the incident that proved
      autonomous agents without cryptographic trust models are dangerous at
      scale --- exactly the failure mode Beadle is designed to prevent.
  \end{enumerate}
\end{faqpair}

% ══════════════════════════════════════════════════════════════════════════════
% FOUR RISKS ASSESSMENT
% ══════════════════════════════════════════════════════════════════════════════

\newpage
\section*{Risk Assessment}

\renewcommand{\arraystretch}{1.6}
\begin{tabularx}{\textwidth}{@{} l l X @{}}
\toprule
\textbf{\color{SectionBlue}Risk} & \textbf{\color{SectionBlue}Rating} & \textbf{\color{SectionBlue}Assessment} \\
\midrule
Value & \textcolor{RiskAmber}{Medium} &
  Developer interest in autonomous agents is proven (Auto-GPT stars,
  LangChain survey). Whether the cryptographic trust model --- signed commands,
  declared permissions, tamperproof audit trail --- resonates with developers
  who already trust their own machines is the core untested assumption.
  The security posture may feel like overhead for personal automation.
  Validation plan: prototype test with 5--10 developers
  (see \faqref{faq:next-step}). \\
Usability & \textcolor{RiskRed}{High} &
  GPG abstraction is the hardest unsolved UX problem in the project. The
  {\small\texttt{beadle init}} command must generate keys and produce an
  identity file without ever exposing raw GPG semantics to the user. SSH
  signing was evaluated and rejected because it lacks encryption and key
  lifecycle primitives (see \faqref{faq:tech-risks}). If the abstraction
  leaks --- if any step requires the user to understand keyrings, trust
  models, or \texttt{gpg} flags --- onboarding friction will kill the
  product. The signing ceremony (sign every command, declare every
  permission) is the trust model's core value but also its primary
  usability risk: developers may perceive it as ceremony, not safety. \\
Feasibility & \textcolor{RiskGreen}{Low} &
  Pipeline v1 is built and e2e verified on Proton Bridge: email arrival
  triggers a multi-stage pipeline (summarize + auto-reply) and delivers
  results to the sender. The trust gate, worker spawner, pipeline executor,
  and crash-safe persistence all work. The remaining integration risk is in
  Pipeline v2's CLI runner and compound step execution --- both are
  well-understood patterns (exec.Command, io.Pipe) with a thorough design
  (3 architect reviews). Proton Bridge has run continuously for weeks as a
  daemon IMAP source without issues. \\
Viability & \textcolor{RiskGreen}{Low} &
  No revenue requirement. Open-source, no infrastructure costs beyond the
  developer's own machine. The only viability risk is opportunity cost ---
  development time spent on Beadle is time not spent on other Punt Labs
  projects. Mitigated by a kill switch: if at 6 months post-release Beadle
  has fewer than 100 GitHub stars or fewer than 10 weekly binary downloads,
  the project enters archival mode and development hours stop
  (see \faqref{faq:pnl} and \faqref{faq:metrics}). \\
\bottomrule
\end{tabularx}
\renewcommand{\arraystretch}{1.0}

% ══════════════════════════════════════════════════════════════════════════════
% FEATURE APPENDIX
% ══════════════════════════════════════════════════════════════════════════════

\newpage
\section*{Feature Appendix}

Defines the scope boundary for the product. Every feature is explicitly
categorized to prevent scope creep and force prioritization decisions upfront.

\subsection*{Shipped}

Features that are built and working.

\begin{enumerate}[nosep,leftmargin=2.5em]
  \featureitem{MCP email server}{17 tools --- list, read, send, move,
    attachments, verify signatures, MIME inspection, trust classification,
    address book, identity switching, inbox polling}\label{feat:mcp}
  \featureitem{Four-level transport trust}{trusted (Proton E2E), verified
    (PGP), untrusted (bad PGP), unverified (none). Inline PGP verification
    during message listing}\label{feat:trust}
  \featureitem{Per-contact rwx permissions}{read, write, execute per contact
    per identity. x-bit gates autonomous pipeline execution}\label{feat:permissions}
  \featureitem{Multi-identity via ethos}{identity resolved per-request from
    ethos sidecar. Repo-local config, global fallback, mid-session
    switching}\label{feat:identity}
  \featureitem{Pipeline v1}{background daemon with IMAP poller, trust gate,
    command YAML loader with typed args, regex planner, sequential executor,
    auto-reply as terminal stage, crash-safe pipeline persistence, worker
    spawner via \texttt{claude -p --bare} with adversarial
    isolation}\label{feat:pipeline-v1}
  \featureitem{Email trigger via Proton Bridge}{receive instructions by email,
    verify sender identity, execute pipeline, return results by
    email}\label{feat:email}
  \featureitem{Credential isolation}{secrets resolved at runtime from OS
    keychain (macOS Keychain, Linux pass/libsecret), secret files, or env
    vars. Never stored in config}\label{feat:credentials}
\end{enumerate}

\subsection*{In Progress (Pipeline v2)}

The next round of work. Designed and architect-reviewed (3 rounds).

\begin{enumerate}[nosep,leftmargin=2.5em]
  \featureitem{CLI runner}{execute tools (biff, jq, git, ethos) directly via
    \texttt{exec.Command} in milliseconds. Whitelist-only binary resolution.
    No shell, no LLM overhead for deterministic
    operations}\label{feat:cli-runner}
  \featureitem{Process/passthrough modes}{commands declare whether they
    transform the pipe (process) or read it as a side-effect (passthrough).
    Notifications don't destroy pipeline data}\label{feat:passthrough}
  \featureitem{Compound CLI steps}{chain binaries within a single command
    (jq | biff) via goroutine-per-step with io.Pipe. No shell
    invoked}\label{feat:compound}
  \featureitem{JSON pipe with schema validation}{structured data flows between
    stages. Commands declare output schemas. Daemon validates JSON between
    process-mode stages}\label{feat:schemas}
  \featureitem{Runner-conditional validation}{prompt and budget required for
    Claude commands only. Binary and steps for CLI only. Typed args for
    both}\label{feat:runner-validation}
\end{enumerate}

\subsection*{Should Do}

Features for after Pipeline v2. Not launch-blocking.

\begin{enumerate}[nosep,leftmargin=2.5em]
  \featureitem{LLM planner}{replace regex planner with Claude decomposition
    of natural language instructions into command
    sequences}\label{feat:llm-planner}
  \featureitem{GPG command signing enforcement}{verify GPG signatures on
    command YAML files at daemon startup. Reject unsigned or tampered
    files}\label{feat:signing}
  \featureitem{Internal scheduler}{cron-like scheduled pipeline execution with
    no catch-up on missed runs}\label{feat:scheduler}
  \featureitem{Multi-channel triggers}{biff chat, webhooks as trigger channels
    alongside email}\label{feat:multi-channel}
  \featureitem{Key rotation management}{guided key rotation with expiry
    enforcement}\label{feat:key-rotation}
  \featureitem{Multi-provider email}{Fastmail, Gmail, self-hosted Dovecot.
    Standard IMAP/SMTP, no architectural lock-in}\label{feat:multi-provider}
\end{enumerate}

\subsection*{Won't Do}

Features explicitly excluded. Naming what you won't build prevents scope
creep and clarifies the product's identity.

\begin{enumerate}[nosep,leftmargin=2.5em]
  \featureitem{Multi-identity support}{Beadle is a personal agent for one
    owner. Multi-tenant adds complexity that conflicts with the single-identity
    trust model}\label{feat:no-multi-identity}
  \featureitem{Signal or other messaging integrations}{email is the remote
    interface. Adding messaging platforms fragments the authentication model
    and increases attack surface}\label{feat:no-signal}
  \featureitem{Web UI}{a browser interface contradicts the local-first,
    CLI-native design philosophy and adds a large attack
    surface}\label{feat:no-web-ui}
  \featureitem{Cloud hosting or SaaS model}{Beadle runs on the owner's machine.
    A hosted version would undermine the core value proposition of data
    sovereignty and local control}\label{feat:no-cloud}
  \featureitem{Scheduler catch-up}{missed scheduled runs are skipped, not
    replayed. Catch-up logic adds complexity and ambiguity about which version
    of a command should execute}\label{feat:no-catchup}
  \featureitem{Non-expiring command-signing keys}{all command-signing keys must
    have an expiration date. This is a security invariant, not a
    convenience trade-off}\label{feat:no-nonexpiring}
\end{enumerate}

% ══════════════════════════════════════════════════════════════════════════════
% REFERENCES
% ══════════════════════════════════════════════════════════════════════════════

\newpage
\printbibliography[title={References}]

\end{document}
