%
% architecture.tex
%
% Beadle Email Channel: Architecture & Z Specification
% Package structure, trust model, credential resolution, message lifecycle,
% and security boundaries for beadle-email v0.1.x.
%
% Type-check: fuzz -t architecture.tex
%

\documentclass[a4paper,10pt,fleqn]{article}
\usepackage[margin=1in]{geometry}
\usepackage{fuzz}
\usepackage[colorlinks=true,linkcolor=blue,citecolor=blue,urlcolor=blue]{hyperref}
\usepackage{booktabs}
\usepackage{longtable}

\begin{document}

\title{Beadle Email Channel: Architecture \& Z Specification}
\author{Punt Labs}
\date{March 2026}
\hypersetup{
  pdftitle={Beadle Email Channel: Architecture and Z Specification},
  pdfauthor={Punt Labs},
  pdfsubject={Architecture Document},
  pdfcreator={fuzz}
}
\maketitle

\tableofcontents
\newpage

%%----------------------------------------------------------------------
\section{Introduction}
%%----------------------------------------------------------------------

Beadle is an autonomous agent daemon with cryptographic owner control.
Every action requires a GPG-signed instruction from the owner; the
daemon has no independent decision-making authority.

The first shipping component is \texttt{beadle-email} --- an MCP server
that provides email communication via Proton Bridge with a four-level
PGP trust model.  This document describes the architecture of
\texttt{beadle-email} v0.1.x: its package structure, data flows,
security boundaries, and key state machines formalised in Z.

%%----------------------------------------------------------------------
\section{System Overview}
%%----------------------------------------------------------------------

\texttt{beadle-email} runs as a stdio-transport MCP server.  It
connects to Proton Bridge via IMAP/STARTTLS on localhost, sends mail
via Proton Bridge SMTP (primary) or Resend API (fallback), and
verifies PGP signatures using an isolated GPG keyring.

The binary exposes four subcommands:

\begin{itemize}
  \item \texttt{serve} --- start the MCP server (default when no
    subcommand is given).
  \item \texttt{version} --- print the version string.
  \item \texttt{doctor} --- check installation health (config,
    credentials, GPG, SMTP reachability).
  \item \texttt{status} --- print current configuration as JSON.
\end{itemize}

%%----------------------------------------------------------------------
\section{Package Map}
%%----------------------------------------------------------------------

All packages live under \texttt{internal/} --- nothing is exported
outside the module.

\begin{center}
\begin{tabular}{@{}ll@{}}
\toprule
\textbf{Package} & \textbf{Responsibility} \\
\midrule
\texttt{cmd/beadle-email/} & CLI entry point: subcommand dispatch \\
\texttt{internal/channel/}  & Channel interface: \texttt{Message}, \texttt{TrustLevel}, shared types \\
\texttt{internal/email/}    & IMAP client, MIME parser, trust classifier, SMTP/Resend senders \\
\texttt{internal/pgp/}      & GPG signature verification and signing via isolated GNUPGHOME \\
\texttt{internal/mcp/}      & MCP tool definitions and handlers (8 tools) \\
\texttt{internal/secret/}   & Credential resolution: OS keychain $\to$ file $\to$ env var \\
\bottomrule
\end{tabular}
\end{center}

\subsection{Dependency Graph}

Dependencies flow strictly downward.  No package imports a package
above it in this ordering.

\begin{verbatim}
    cmd/beadle-email
          |
      internal/mcp
       /    |    \
  email    pgp   channel
    |
  secret
\end{verbatim}

\texttt{internal/channel} is a leaf package with no internal
dependencies.  \texttt{internal/pgp} depends only on the standard
library.  \texttt{internal/email} depends on \texttt{channel} and
\texttt{secret}.  \texttt{internal/mcp} depends on \texttt{email},
\texttt{pgp}, and \texttt{channel}.

%%----------------------------------------------------------------------
\section{Basic Types}
%%----------------------------------------------------------------------

\begin{zed}
  [MSGID, ADDRESS, BODY, HEADER, RAWBYTES, KEYID, SIGNER, CREDNAME]
\end{zed}

$MSGID$ is an IMAP UID (numeric string).  $ADDRESS$ is an RFC~5322
email address.  $BODY$ is the plain-text message body.  $HEADER$ is
a header name-value pair.  $RAWBYTES$ is the raw RFC~822 message.
$KEYID$ is a GPG key fingerprint.  $SIGNER$ is a GPG user identity.
$CREDNAME$ is a named credential (e.g.\ \texttt{imap-password}).

%%----------------------------------------------------------------------
\section{Free Types}
%%----------------------------------------------------------------------

\begin{zed}
  ZBOOL ::= ztrue | zfalse
\end{zed}

\begin{zed}
  TrustLevel ::= trusted | verified | untrusted | unverified
\end{zed}

Four trust levels classify inbound messages:

\begin{itemize}
  \item $trusted$ --- Proton-to-Proton, end-to-end encrypted by Proton
    infrastructure.
  \item $verified$ --- external sender with a valid PGP signature.
  \item $untrusted$ --- external sender with an invalid PGP signature.
  \item $unverified$ --- external sender with no signature present.
\end{itemize}

\begin{zed}
  SendMethod ::= smProtonSMTP | smResend
\end{zed}

Outbound messages use Proton Bridge SMTP as the primary transport,
falling back to Resend API when Bridge is unavailable.

\begin{zed}
  CredBackend ::= cbKeychain | cbFile | cbEnvVar
\end{zed}

Credentials are resolved through a priority chain of three backends.

\begin{zed}
  IMAPState ::= imapDisconnected | imapConnected | imapSelected
\end{zed}

The IMAP client progresses through connection, authentication, and
folder selection before fetching messages.

%%----------------------------------------------------------------------
\section{Trust Model}
%%----------------------------------------------------------------------

This section formalises the trust classification state machine.
Trust classification is a pure function of message headers and
PGP verification outcome.

\begin{schema}{TrustInput}
  hasProtonE2E : ZBOOL \\
  hasProtonInternal : ZBOOL \\
  hasPGPSignature : ZBOOL \\
  pgpVerified : ZBOOL
\end{schema}

$hasProtonE2E$ is true when the header
\texttt{X-Pm-Content-Encryption} contains \texttt{end-to-end}.
$hasProtonInternal$ is true when \texttt{X-Pm-Origin} is
\texttt{internal}.  $hasPGPSignature$ is true when the Content-Type
is \texttt{multipart/signed} or raw bytes contain
\texttt{application/pgp-signature}.  $pgpVerified$ is the result of
\texttt{gpg --verify} (only meaningful when $hasPGPSignature = ztrue$).

\begin{schema}{ClassifyTrust}
  TrustInput \\
  level! : TrustLevel
  \where
  (hasProtonE2E = ztrue \land hasProtonInternal = ztrue
    \implies level! = trusted) \\
  \land \\
  (hasPGPSignature = ztrue \land pgpVerified = ztrue
    \land \lnot (hasProtonE2E = ztrue \land hasProtonInternal = ztrue)
    \implies level! = verified) \\
  \land \\
  (hasPGPSignature = ztrue \land pgpVerified = zfalse
    \land \lnot (hasProtonE2E = ztrue \land hasProtonInternal = ztrue)
    \implies level! = untrusted) \\
  \land \\
  (hasPGPSignature = zfalse
    \land \lnot (hasProtonE2E = ztrue \land hasProtonInternal = ztrue)
    \implies level! = unverified)
\end{schema}

The classification is total: every combination of inputs maps to
exactly one trust level.  Proton E2E takes priority over PGP
signatures (a Proton-internal message with a PGP signature is
still $trusted$, not $verified$).

\subsection{Two-Phase Trust Resolution}

Trust classification happens in two phases because PGP verification
requires fetching the raw message bytes and running \texttt{gpg}:

\begin{enumerate}
  \item \textbf{Header scan} --- \texttt{ClassifyTrust} runs with
    $pgpVerified = zfalse$.  Messages with signatures return
    $unverified$ with $hasPGPSignature = ztrue$, signalling the
    caller to proceed to phase~2.
  \item \textbf{PGP verification} --- the caller fetches raw bytes,
    runs \texttt{gpg --verify} in an isolated GNUPGHOME, and
    re-classifies with the actual $pgpVerified$ value.
\end{enumerate}

%%----------------------------------------------------------------------
\section{Credential Resolution}
%%----------------------------------------------------------------------

\begin{schema}{CredentialStore}
  keychainAvail : ZBOOL \\
  keychainValues : CREDNAME \pfun BODY \\
  fileValues : CREDNAME \pfun BODY \\
  envValues : CREDNAME \pfun BODY
\end{schema}

Credentials are resolved by name.  The keychain backend maps to
OS-specific stores.  On Darwin, the macOS Keychain via the
\texttt{security} CLI, keyed by (service=\texttt{beadle},
account=\texttt{<name>}).  On Linux, a two-stage lookup: first
\texttt{pass} (\texttt{pass show beadle/<name>}), which keeps
secrets GPG-encrypted with the user's own key and matches the
Proton Bridge vault backend; then \texttt{secret-tool} (libsecret
/ GNOME Keyring), keyed by the same (service, account) pair as
Darwin for portability.  If both Linux binaries are present,
\texttt{pass} wins.  File credentials live in
\texttt{\~{}/.punt-labs/beadle/secrets/<name>} with mode 600.
Environment variables follow the pattern \texttt{BEADLE\_<NAME>}.

\begin{schema}{ResolveCredential}
  \Xi CredentialStore \\
  name? : CREDNAME \\
  value! : BODY
  \where
  (keychainAvail = ztrue \land name? \in \dom keychainValues
    \implies value! = keychainValues(name?)) \\
  \land \\
  ((\lnot (keychainAvail = ztrue \land name? \in \dom keychainValues))
    \land name? \in \dom fileValues
    \implies value! = fileValues(name?)) \\
  \land \\
  ((\lnot (keychainAvail = ztrue \land name? \in \dom keychainValues))
    \land name? \notin \dom fileValues
    \land name? \in \dom envValues
    \implies value! = envValues(name?))
\end{schema}

The resolution is a priority chain: keychain beats file, file beats
environment variable.  File credentials are rejected if group or
world readable (permissions check enforces mode 600).

The three named credentials used by \texttt{beadle-email}:

\begin{center}
\begin{tabular}{@{}lll@{}}
\toprule
\textbf{Name} & \textbf{Env Var} & \textbf{Purpose} \\
\midrule
\texttt{imap-password}   & \texttt{BEADLE\_IMAP\_PASSWORD}   & Proton Bridge login \\
\texttt{resend-api-key}  & \texttt{BEADLE\_RESEND\_API\_KEY}  & Resend API authentication \\
\texttt{gpg-passphrase}  & \texttt{BEADLE\_GPG\_PASSPHRASE}  & PGP signing key passphrase \\
\bottomrule
\end{tabular}
\end{center}

%%----------------------------------------------------------------------
\section{Message Receive Path}
%%----------------------------------------------------------------------

The receive path is the data flow from IMAP mailbox to MCP tool
response.

\begin{schema}{IMAPConnection}
  state : IMAPState \\
  host : ADDRESS \\
  user : ADDRESS
\end{schema}

\begin{schema}{Connect}
  \Delta IMAPConnection
  \where
  state = imapDisconnected \\
  state' = imapConnected
\end{schema}

Connection uses STARTTLS.  The TLS certificate check is skipped for
loopback addresses only (Proton Bridge uses self-signed certificates
on localhost).

\begin{schema}{SelectFolder}
  \Delta IMAPConnection \\
  folder? : CREDNAME
  \where
  state = imapConnected \lor state = imapSelected \\
  state' = imapSelected
\end{schema}

\subsection{Fetch and Parse}

Once a folder is selected, messages are fetched by UID.  The raw
RFC~822 bytes are parsed through a pipeline:

\begin{enumerate}
  \item \textbf{MIME parse} --- \texttt{ParseMIME} extracts the
    plain-text body, attachments, and selected headers.  Multipart
    messages are walked using the \texttt{go-message} library.
    PGP signature and key parts are skipped (handled separately).
  \item \textbf{Header trust} --- \texttt{ClassifyTrust} runs on
    headers to determine the initial trust level.
  \item \textbf{PGP verify} (conditional) --- if the message has a
    PGP signature ($hasPGPSignature = ztrue$), raw bytes are
    passed to \texttt{pgp.Verify} for full verification.
  \item \textbf{Response assembly} --- the parsed message, trust
    level, and attachments are serialised as JSON in the MCP tool
    response.
\end{enumerate}

%%----------------------------------------------------------------------
\section{Message Send Path}
%%----------------------------------------------------------------------

\begin{schema}{SendState}
  smtpAvailable : ZBOOL \\
  method! : SendMethod
  \where
  smtpAvailable = ztrue \implies method! = smProtonSMTP \\
  smtpAvailable = zfalse \implies method! = smResend
\end{schema}

The send path attempts Proton Bridge SMTP first.  If the Bridge is
unreachable (a TCP dial to the SMTP port times out in 2~seconds),
the Resend API is used as fallback.

\subsection{Proton Bridge SMTP}

SMTP uses STARTTLS with the same self-signed certificate exception
as IMAP (loopback only).  Authentication uses \texttt{PLAIN} with
the same credentials as IMAP.  The raw RFC~822 message is composed
by \texttt{ComposeRaw}, which validates header fields against CR/LF
injection.

\subsection{Resend API}

The Resend fallback sends via HTTPS POST to
\texttt{api.resend.com/emails}.  It supports both plain text and
HTML bodies.  The API key is resolved via the credential store.

\subsection{PGP Signing (Future)}

PGP signing of outbound mail requires a transport that preserves raw
MIME (\texttt{multipart/signed} envelopes).  Neither Proton Bridge
(strips MIME) nor Resend (no raw MIME API) supports this.  Amazon
SES will be added as the PGP-signing transport in a future release.

%%----------------------------------------------------------------------
\section{PGP Subsystem}
%%----------------------------------------------------------------------

The PGP subsystem handles signature verification and signing via
the system \texttt{gpg} binary.  It never uses a Go PGP library ---
\texttt{gpg} handles keyring management, trust models, and
signature formats.

\begin{schema}{GPGEnvironment}
  systemKeys : \power KEYID \\
  importedKeys : \power KEYID \\
  isolated : ZBOOL
  \where
  isolated = ztrue \\
  importedKeys \subseteq systemKeys
\end{schema}

Every verification runs in an isolated GNUPGHOME under
\texttt{/tmp/bg-*}.  The short path prefix avoids the 108-byte
Unix socket path limit for \texttt{gpg-agent}.

\subsection{Verify Flow}

\begin{enumerate}
  \item Extract signed data, detached signature, and optional public
    key from the \texttt{multipart/signed} envelope (RFC~3156).
  \item Create an isolated GNUPGHOME directory.
  \item If the message includes a public key attachment
    (\texttt{application/pgp-keys}), import it into the isolated
    keyring.
  \item If no key was attached, export all public keys from the
    system keyring (\texttt{\~{}/.gnupg}) into the isolated keyring.
    This is a one-way bridge: read from system, never write.
  \item Run \texttt{gpg --verify} with the detached signature and
    signed data.
  \item Parse the exit code and stderr for validity, signer identity,
    and key ID.
  \item Remove the temporary directory.
\end{enumerate}

\subsection{Sign Flow}

\begin{enumerate}
  \item Compose the body as a MIME part with CRLF line endings
    (mandatory per RFC~3156).
  \item Write the GPG passphrase to a temporary file (mode 600) to
    avoid exposing it in process arguments or \texttt{ps} output.
  \item Run \texttt{gpg --detach-sign --armor} with
    \texttt{--passphrase-file}.
  \item Assemble the full RFC~822 message with a
    \texttt{multipart/signed} envelope containing the body part and
    detached signature.
  \item Generate the MIME boundary using \texttt{crypto/rand}.
\end{enumerate}

%%----------------------------------------------------------------------
\section{MCP Tool Surface}
%%----------------------------------------------------------------------

\texttt{beadle-email} registers 8 MCP tools:

\begin{center}
\begin{longtable}{@{}llp{6.5cm}@{}}
\toprule
\textbf{Tool} & \textbf{Required Params} & \textbf{Description} \\
\midrule
\texttt{list\_messages}     & ---               & List messages from a folder (id, from, date, subject, trust). Inline PGP verification for signed messages. \\
\texttt{read\_message}      & \texttt{message\_id} & Read a single message by UID. Auto-verifies PGP if Content-Type indicates \texttt{multipart/signed}. \\
\texttt{list\_folders}      & ---               & List all IMAP mailbox folders. \\
\texttt{send\_email}        & \texttt{to}, \texttt{subject}, \texttt{body} & Send via SMTP (primary) or Resend (fallback). \\
\texttt{verify\_signature}  & \texttt{message\_id} & Explicit PGP verification. Returns trust level, validity, key ID, signer. \\
\texttt{show\_mime}         & \texttt{message\_id} & Show MIME structure of a message. \\
\texttt{check\_trust}       & \texttt{message\_id} & Classify trust level with detailed explanation. \\
\texttt{move\_message}      & \texttt{message\_id} & Move a message to another folder (default: Archive). \\
\bottomrule
\end{longtable}
\end{center}

All tool handlers use a \texttt{withClient} pattern: a fresh IMAP
connection is established per request and closed when the handler
returns.  This avoids stale connection state at the cost of one
TCP handshake per tool call.

%%----------------------------------------------------------------------
\section{Design Invariants}
%%----------------------------------------------------------------------

These invariants are enforced by the architecture and hold across
all operations:

\begin{enumerate}
  \item \textbf{Zero agent authority.}  Every action requires a
    GPG-signed instruction from the owner.  The daemon has no
    independent decision-making.

  \item \textbf{Preflight before execute.}  All permissions are
    validated before any command runs.  No partial execution.

  \item \textbf{Isolated keychain.}  PGP operations use temporary
    GNUPGHOME directories, never touching the user's system GPG
    keyring for writes.

  \item \textbf{Non-expiring keys rejected.}  All command-signing
    keys must have an expiration date.

  \item \textbf{Audit log is tamperproof.}  Append-only,
    GPG-signed entries.  Only the owner can clear the log.

  \item \textbf{No secrets in logs.}  GPG key material, passwords,
    API keys, and raw email content are never logged.

  \item \textbf{No shell execution.}  All \texttt{exec.Command}
    calls use argument lists, never \texttt{shell=true}.

  \item \textbf{Header injection prevention.}  \texttt{ComposeRaw}
    and \texttt{Sign} reject header fields containing CR/LF.

  \item \textbf{File permission enforcement.}  Secret files must
    not be group or world readable (mode 600).

  \item \textbf{Loopback-only TLS exception.}  Self-signed
    certificate verification is skipped only for loopback
    addresses (Proton Bridge on localhost).
\end{enumerate}

%%----------------------------------------------------------------------
\section{Security Boundaries}
%%----------------------------------------------------------------------

\subsection{Network Boundary}

\begin{itemize}
  \item \textbf{IMAP/SMTP}: localhost only (Proton Bridge).
    STARTTLS with self-signed cert exception for loopback.
  \item \textbf{Resend API}: outbound HTTPS to
    \texttt{api.resend.com}.  Bearer token authentication.
  \item \textbf{MCP}: stdio transport.  No network listener.
\end{itemize}

\subsection{Process Boundary}

\begin{itemize}
  \item \textbf{GPG}: subprocess with isolated GNUPGHOME.
    Passphrase passed via temp file (mode 600), never via
    arguments or environment.
  \item \textbf{Keychain}: subprocess (\texttt{security} on macOS,
    \texttt{secret-tool} on Linux).  No direct library binding.
\end{itemize}

\subsection{Trust Boundary}

\begin{itemize}
  \item \textbf{Trusted}: Proton-to-Proton messages.  E2E
    encryption verified by Proton infrastructure, not by beadle.
  \item \textbf{Verified}: external messages with valid PGP
    signatures.  Verification performed by beadle in isolated GPG.
  \item \textbf{Untrusted}: external messages with invalid PGP
    signatures.  The signature was present but did not verify.
  \item \textbf{Unverified}: external messages with no signature.
    No cryptographic assertion about the sender.
\end{itemize}

%%----------------------------------------------------------------------
\section{Key Properties}
%%----------------------------------------------------------------------

The specification captures these properties, verifiable by
inspection and testing:

\begin{enumerate}
  \item \textbf{Trust totality}: every inbound message receives
    exactly one of the four trust levels.  The
    \texttt{ClassifyTrust} schema is total over all input
    combinations.

  \item \textbf{Credential priority}: keychain always takes
    priority over file, file over environment variable.  The
    \texttt{ResolveCredential} schema enforces this ordering.

  \item \textbf{GPG isolation}: the $isolated = ztrue$ invariant
    in \texttt{GPGEnvironment} ensures no PGP operation writes to
    the user's system keyring.

  \item \textbf{Send fallback}: if SMTP is available, it is
    always preferred.  Resend is used only when SMTP is
    unreachable.

  \item \textbf{Connection-per-request}: each MCP tool call
    creates and destroys its own IMAP connection, preventing
    stale state across calls.
\end{enumerate}

\end{document}
