%
% audit-beadle.tex
%
% Beadle audit against Punt Labs user-facing standards
% Generated 2026-03-18
%
% Compile: pdflatex audit-beadle.tex
%

\documentclass[a4paper,10pt]{article}
\usepackage[margin=0.8in]{geometry}
\usepackage{enumitem}
\usepackage{booktabs}
\usepackage[colorlinks=true,linkcolor=blue]{hyperref}
\usepackage{fancyhdr}
\usepackage{titlesec}
\usepackage{amssymb}
\usepackage{xcolor}

\titlespacing*{\section}{0pt}{2ex}{0.5ex}
\titlespacing*{\subsection}{0pt}{1.2ex}{0.3ex}

\newcommand{\pass}{{\color{green!60!black}$\blacksquare$\hspace{0.4em}\textsc{pass}}\hspace{0.5em}}
\newcommand{\fail}{{\color{red!70!black}$\blacksquare$\hspace{0.4em}\textsc{fail}}\hspace{0.5em}}
\newcommand{\na}{{\color{gray}$\square$\hspace{0.4em}\textsc{n/a}}\hspace{0.5em}}

\setlist[itemize]{nosep, left=0pt, labelsep=0em, label={}}

\pagestyle{fancy}
\fancyhf{}
\rhead{Beadle Audit --- 2026-03-18}
\lhead{\leftmark}
\rfoot{\thepage}

\begin{document}

\title{Beadle: User-Facing Standards Audit}
\author{Project: \texttt{punt-labs/beadle} (v0.3.1)}
\date{March 18, 2026}
\maketitle

\begin{center}
\begin{tabular}{lrrr}
\toprule
& \textsc{pass} & \textsc{fail} & \textsc{n/a} \\
\midrule
Naming \& Identity    & 2 & 2 & 1 \\
CLI Design            & 8 & 7 & 0 \\
Permissions \& Safety & 8 & 1 & 0 \\
Plugin Experience     & 9 & 0 & 0 \\
Hooks                 & 7 & 2 & 2 \\
Installation          & 3 & 5 & 0 \\
Filesystem Layout     & 0 & 4 & 1 \\
README \& Docs        & 2 & 1 & 1 \\
Integration           & 2 & 0 & 3 \\
\midrule
\textbf{Total}        & \textbf{41} & \textbf{22} & \textbf{8} \\
\bottomrule
\end{tabular}
\end{center}

\tableofcontents
\newpage

% ══════════════════════════════════════════════════════════════════════
\section{Naming \& Identity}

\begin{itemize}
  \item \fail Project name = CLI command = repo name ---
    Repo is \texttt{beadle}, CLI is \texttt{beadle-email}. Mismatch.
  \item \na PyPI package --- Go project, not applicable.
  \item \pass No \texttt{-mcp} suffix.
  \item \pass Slash commands: \texttt{/inbox}, \texttt{/mail}, \texttt{/send}.
  \item \fail Version match --- \texttt{plugin.json} has 0.3.1,
    \texttt{main.go} uses ldflags (\texttt{dev} default),
    \texttt{install.sh} has no VERSION pin.
\end{itemize}

% ══════════════════════════════════════════════════════════════════════
\section{CLI Design}

\subsection{Required Subcommands}
\begin{itemize}
  \item \pass \texttt{version}
  \item \pass \texttt{doctor} --- per-dependency pass/fail checks
  \item \pass \texttt{status} --- JSON state summary
  \item \fail \texttt{install} --- missing; no machine-scoped setup command
  \item \pass \texttt{serve} --- stdio transport
\end{itemize}

\subsection{Subcommand Naming}
\begin{itemize}
  \item \pass Single verbs for all subcommands
  \item \pass Arguments disambiguate
  \item \fail Product commands first in \texttt{--help} --- admin commands
    (\texttt{serve}) listed before product commands (\texttt{contact})
\end{itemize}

\subsection{Global Flags}
\begin{itemize}
  \item \fail \texttt{--json} --- not implemented
  \item \fail \texttt{--verbose} --- uses \texttt{BEADLE\_DEBUG} env var instead
  \item \fail \texttt{--quiet} --- not implemented
\end{itemize}

\subsection{CLI $\leftrightarrow$ MCP Parity}
\begin{itemize}
  \item \fail Core email operations (list, read, send, move) exist only as
    MCP tools. No CLI subcommands for email operations.
  \item \pass Contact operations have full CLI/MCP parity.
  \item \pass Slash commands map to MCP tools.
  \item \fail Hooks call shell scripts directly, not
    \texttt{beadle-email hook <event>} dispatcher.
\end{itemize}

% ══════════════════════════════════════════════════════════════════════
\section{Permissions \& Safety}

\begin{itemize}
  \item \pass \texttt{settings.json} portable, checked in
  \item \pass \texttt{settings.local.json} gitignored
  \item \pass MCP wildcards auto-allowed by SessionStart
  \item \pass Skill permissions auto-allowed
  \item \pass All required deny rules present (destructive, network, secrets)
  \item \pass Permission injection via \texttt{jq} atomic merge
  \item \pass Pattern-specific rules (include plugin name)
  \item \fail No uninstaller to remove injected permissions
\end{itemize}

% ══════════════════════════════════════════════════════════════════════
\section{Plugin Experience}

\begin{itemize}
  \item \pass Zero-prompt install --- SessionStart auto-allows everything
  \item \pass \texttt{plugin.json}: name, description, version, author all correct
  \item \pass Dev/prod isolation: \texttt{beadle-dev} / \texttt{beadle}
  \item \pass Command deployment: diff-and-copy, skips \texttt{-dev.md}
  \item \pass Output suppression: all 13 MCP tools have handlers
  \item \pass Two-channel display working
  \item \pass Matcher covers dev and prod namespaces
  \item \pass \texttt{allowed-tools} frontmatter correct
  \item \pass MCP-first commands
\end{itemize}

% ══════════════════════════════════════════════════════════════════════
\section{Hooks}

\begin{itemize}
  \item \fail Three-layer dispatch --- shell scripts contain business logic
    directly (session-start: 182 lines, suppress-output: 164 lines).
    Standard requires shell thin gate $\to$ CLI handler.
  \item \fail Business logic in shell exceeds 10-line threshold.
  \item \pass Observation hooks fail-open (both exit 0 on all paths)
  \item \pass Matcher covers dev and prod
  \item \pass SessionStart: deploys commands, auto-allows, first-run check
  \item \pass PostToolUse: all 13 tools suppressed with two-channel display
  \item \pass No dead code, no orphan handlers
\end{itemize}

% ══════════════════════════════════════════════════════════════════════
\section{Installation \& Distribution}

\begin{itemize}
  \item \pass One-command install via \texttt{curl | sh}
  \item \pass POSIX \texttt{install.sh}, idempotent, transparent
  \item \pass Stdin protection on claude/ssh commands
  \item \fail URL pinned to SHA in README but \texttt{install.sh} itself
    references \texttt{main} in a comment
  \item \fail No VERSION pin in \texttt{install.sh} --- downloads
    \texttt{releases/latest}
  \item \fail Missing ``verify'' trust tier (only convenience + inspect)
  \item \fail \texttt{install.sh} tells user to run \texttt{doctor} but
    does not run it automatically
  \item \fail No uninstall mechanism (script or CLI command)
\end{itemize}

% ══════════════════════════════════════════════════════════════════════
\section{Filesystem Layout}

\begin{itemize}
  \item \fail Three root paths instead of one:
    \texttt{\textasciitilde/.punt-labs/beadle/} (contacts),
    \texttt{\textasciitilde/.config/beadle/} (config),
    \texttt{\textasciitilde/.beadle/} (attachments)
  \item \fail \texttt{\textasciitilde/.beadle/} and
    \texttt{\textasciitilde/.config/beadle/} are non-compliant top-level dirs
  \item \fail No single root constant --- three separate path derivations
  \item \na Per-project sentinel --- not applicable (user-level tool)
  \item \fail No file-based logging under
    \texttt{\textasciitilde/.punt-labs/beadle/logs/}
\end{itemize}

% ══════════════════════════════════════════════════════════════════════
\section{README \& Documentation}

\begin{itemize}
  \item \pass All required sections present in correct order
  \item \pass Tone: factual, no marketing, no anthropomorphizing
  \item \fail License section says ``TBD'' instead of MIT
  \item \na Competitor mentions --- not needed
\end{itemize}

% ══════════════════════════════════════════════════════════════════════
\section{Integration}

\begin{itemize}
  \item \pass Graceful degradation (SMTP $\to$ Resend fallback, missing
    contacts file not an error)
  \item \pass Sole writer (beadle only writes to its own paths)
  \item \na L0/L1 discovery --- no inter-tool checks
\end{itemize}

% ══════════════════════════════════════════════════════════════════════
\newpage
\section{Priority Fixes for Release Readiness}

Ordered by user impact:

\begin{enumerate}
  \item \textbf{Filesystem consolidation} --- migrate
    \texttt{\textasciitilde/.config/beadle/} and
    \texttt{\textasciitilde/.beadle/} to
    \texttt{\textasciitilde/.punt-labs/beadle/}. Single root constant.
  \item \textbf{CLI parity for email ops} --- add \texttt{list}, \texttt{read},
    \texttt{send}, \texttt{move} CLI subcommands calling same internal functions.
  \item \textbf{install subcommand} --- \texttt{beadle-email install} for
    MCP registration, data dir creation, credential prompting.
  \item \textbf{uninstall subcommand} --- remove deployed commands, permissions,
    MCP registration.
  \item \textbf{VERSION pin in install.sh} --- pin to specific version, not
    \texttt{latest}.
  \item \textbf{License} --- set to MIT in README and add LICENSE file.
  \item \textbf{Global flags} --- \texttt{--json}, \texttt{--verbose},
    \texttt{--quiet}.
  \item \textbf{Hook refactor} --- extract shell business logic to Go CLI
    handler (\texttt{beadle-email hook session-start}).
  \item \textbf{Version consistency} --- single source of truth across
    plugin.json, main.go, install.sh.
  \item \textbf{install.sh doctor} --- auto-run \texttt{doctor} at end.
\end{enumerate}

\end{document}
