%
% checklist.tex
%
% Punt Labs Standards Audit Checklist
% Derived from punt-kit/standards/*.md
%
% Compile: pdflatex checklist.tex && pdflatex checklist.tex
%

\documentclass[a4paper,10pt]{article}
\usepackage[margin=0.8in]{geometry}
\usepackage{enumitem}
\usepackage{booktabs}
\usepackage[colorlinks=true,linkcolor=blue,citecolor=blue,urlcolor=blue]{hyperref}
\usepackage{fancyhdr}
\usepackage{titlesec}
\usepackage{amssymb}

\titlespacing*{\section}{0pt}{2ex}{0.5ex}
\titlespacing*{\subsection}{0pt}{1.2ex}{0.3ex}

\newcommand{\must}{\textsc{must}}
\newcommand{\should}{\textsc{should}}

\setlist[itemize]{nosep, left=1.5em, labelsep=0.5em, label={$\square$}}
\setlist[enumerate]{nosep}

\pagestyle{fancy}
\fancyhf{}
\rhead{Punt Labs Standards Audit}
\lhead{\leftmark}
\rfoot{\thepage}

\begin{document}

\title{Punt Labs Standards Audit Checklist}
\author{Derived from \texttt{punt-kit/standards/*.md}}
\date{\today}
\maketitle

\noindent Two parts: \textbf{Part~I} covers what users see and touch ---
naming, CLI, permissions, hooks, installation, documentation.
\textbf{Part~II} covers developer workflow --- git, CI, build, code
standards, release.

\tableofcontents
\newpage

% 
%   PART I: USER-FACING                                             
% 

\part{User-Facing}

% ──────────────────────────────────────────────────────────────────────
\section{Naming \& Identity}
\label{sec:naming}
Sources: \texttt{naming.md}, \texttt{cli.md~\S3}

\begin{itemize}
  \item Project name = CLI command = GitHub repo name
  \item PyPI package: \texttt{punt-<name>}; CLI entry point: short name, no prefix
  \item No \texttt{-mcp} suffix --- the CLI \emph{is} the product, MCP is a projection
  \item Slash commands: \texttt{/<name>} or \texttt{/<name>:<sub>}
  \item Version matches across all locations (pyproject.toml, plugin.json, install.sh, \_\_init\_\_.py)
\end{itemize}

% ──────────────────────────────────────────────────────────────────────
\section{CLI Design}
\label{sec:cli}
Source: \texttt{cli.md}

\subsection{Core Principle}
The CLI is the complete product. Every capability must be accessible from the
terminal. MCP tools, slash commands, and hooks are projections --- a user
without Claude Code can use every feature.

\subsection{Required Subcommands (\must)}
\begin{itemize}
  \item \texttt{version} --- print version
  \item \texttt{doctor} --- per-dependency pass/fail health check
  \item \texttt{status} --- current state summary
  \item \texttt{install} --- machine-scoped setup (MCP registration, data dirs)
  \item \texttt{serve} --- start MCP server (stdio transport)
\end{itemize}

\subsection{Subcommand Naming (\must)}
\begin{itemize}
  \item Single verbs: \texttt{search}, \texttt{ingest}, \texttt{talk}
  \item Arguments disambiguate, not subcommand names
  \item Product commands first in \texttt{--help}, admin commands after
\end{itemize}

\subsection{Global Flags (\should)}
\begin{itemize}
  \item \texttt{--json} / \texttt{-j}: JSON output for every command
  \item \texttt{--verbose} / \texttt{-v}: debug logging
  \item \texttt{--quiet} / \texttt{-q}: errors only
\end{itemize}

\subsection{CLI $\leftrightarrow$ MCP Parity (\must)}
\begin{itemize}
  \item New capabilities implemented in CLI first
  \item MCP tools call same functions as CLI
  \item Slash commands map to CLI commands via MCP tools
  \item Hooks call \texttt{<tool> hook <event>} dispatcher
\end{itemize}

% ──────────────────────────────────────────────────────────────────────
\section{Permissions \& Safety}
\label{sec:permissions}
Source: \texttt{permissions.md}

\subsection{Settings File Split (\must)}
\begin{itemize}
  \item \texttt{settings.json}: portable permissions (checked in)
  \item \texttt{settings.local.json}: machine-specific (gitignored)
\end{itemize}

\subsection{Auto-Allow Rules (\must)}
\begin{itemize}
  \item MCP plugin wildcards: \texttt{mcp\_\_plugin\_*}
  \item Skill permissions: \texttt{Skill(<name>)} for every plugin skill
  \item SessionStart hook populates both automatically
\end{itemize}

\subsection{Required Deny Rules (\must)}
\begin{itemize}
  \item Destructive: \texttt{rm -rf /}, \texttt{rm -rf \textasciitilde}, \texttt{dd}, \texttt{sudo}, \texttt{su}
  \item Network: curl, wget, ssh, scp, ftp, nc, telnet, socat
  \item Secrets: \texttt{Edit(.env)}, \texttt{Write(.env)}, \texttt{Edit(.envrc)}, \texttt{Write(.envrc)}
\end{itemize}

\subsection{Plugin-Distributed Permissions (\must)}
\begin{itemize}
  \item Installer injects permissions via \texttt{jq} atomic merge
  \item Uninstaller removes injected permissions
  \item Pattern-specific (include plugin name in rules)
\end{itemize}

% ──────────────────────────────────────────────────────────────────────
\section{Plugin Experience}
\label{sec:plugins}
Source: \texttt{plugins.md}

\subsection{Zero-Prompt Install (\must)}
Users should never see a permission prompt after enabling the plugin.
SessionStart auto-allows all MCP tools and skills.

\subsection{plugin.json (\must)}
\begin{itemize}
  \item \texttt{name}, \texttt{description}, \texttt{version} (semver)
  \item \texttt{author}: ``Punt Labs'', email: \texttt{hello@punt-labs.com}
  \item Dev name: \texttt{<project>-dev}; prod name: \texttt{<project>}
\end{itemize}

\subsection{Command Deployment (\must)}
\begin{itemize}
  \item SessionStart copies commands to \texttt{\textasciitilde/.claude/commands/}
  \item Diff-and-copy on every session (not skip-if-exists)
  \item Skips \texttt{-dev.md} files in prod mode
\end{itemize}

\subsection{Output Suppression (\must)}
\begin{itemize}
  \item PostToolUse hook suppresses raw MCP JSON
  \item Every MCP tool has a handler in suppress-output
  \item Two-channel display: compact panel summary + full data in context
  \item Matcher covers dev and prod namespaces
\end{itemize}

\subsection{Command Frontmatter (\must)}
\begin{itemize}
  \item \texttt{allowed-tools} lists only tools actually used
  \item MCP-first: prefer MCP tools over Bash/Read/Write
\end{itemize}

% ──────────────────────────────────────────────────────────────────────
\section{Hooks}
\label{sec:hooks}
Source: \texttt{hooks.md}

\subsection{Architecture (\must)}
Three layers: \texttt{hooks.json} registration $\to$ shell thin gate $\to$
CLI/language handler. No business logic in shell scripts ($<$10 lines
non-boilerplate).

\subsection{Behavior (\must)}
\begin{itemize}
  \item Observation hooks fail-open (\texttt{|| true})
  \item Mutation hooks fail-closed (propagate errors)
  \item Matcher regex covers dev and prod: \texttt{mcp\_\_plugin\_<t>(-dev)?\_<s>\_\_.*}
  \item \texttt{async: true} for fire-and-forget hooks
\end{itemize}

\subsection{Required Hooks (marketplace plugins)}
\begin{itemize}
  \item \textbf{SessionStart}: deploy commands, auto-allow permissions, first-run setup
  \item \textbf{PostToolUse}: suppress output, two-channel display for all MCP tools
\end{itemize}

\subsection{Audit}
\begin{itemize}
  \item Every shell script calls its handler (no dead code)
  \item Every handler has \texttt{hooks.json} registration (no orphans)
  \item No \texttt{sys.stdin.read()} in Python handlers
\end{itemize}

% ──────────────────────────────────────────────────────────────────────
\section{Installation \& Distribution}
\label{sec:distribution}
Source: \texttt{distribution.md}

\subsection{Install Experience (\must)}
\begin{itemize}
  \item One command: \texttt{curl -fsSL .../SHA/install.sh | sh}
  \item URL pinned to commit SHA (not \texttt{main})
  \item Three trust tiers in README: convenience, inspect, verify
  \item Ends with \texttt{doctor} to confirm success
\end{itemize}

\subsection{install.sh (\must)}
\begin{itemize}
  \item POSIX: \texttt{\#!/bin/sh}, shellcheck with \texttt{--shell=sh}
  \item Idempotent and transparent (prints each step)
  \item Stdin protection: \texttt{< /dev/null} on all claude/ssh commands
  \item VERSION and EXTRAS pins (mandatory)
\end{itemize}

\subsection{Uninstall (\must)}
\begin{itemize}
  \item Removes deployed commands, MCP permissions, status line wrapping
  \item Removes marketplace registration (conditionally)
\end{itemize}

% ──────────────────────────────────────────────────────────────────────
\section{Filesystem Layout}
\label{sec:filesystem}
Source: \texttt{filesystem.md}

\begin{itemize}
  \item All tool data under \texttt{\textasciitilde/.punt-labs/<tool>/}
  \item No per-tool top-level dot-directories in \texttt{\$HOME}
  \item Single root constant in code; all paths derive from it
  \item Per-project sentinel files at repo root (e.g., \texttt{.biff}, \texttt{.beads/})
  \item Logs: \texttt{\textasciitilde/.punt-labs/<tool>/logs/} (5\,MB rotation, 5 backups)
\end{itemize}

% ──────────────────────────────────────────────────────────────────────
\section{README \& Documentation}
\label{sec:readme}
Source: \texttt{readme.md}

\subsection{Required Sections (in order)}
Title + tagline, Badges (max 4), Description (1 paragraph), Quick Start,
Commands/API table, Development, License (MIT).

\subsection{Tone}
Describe, don't sell. No marketing verbs (``Unlock'', ``Supercharge'').
No anthropomorphizing (``understands'', ``thinks'').
Respect competitors --- name them, link them, praise what they do well.

% ──────────────────────────────────────────────────────────────────────
\section{Integration}
\label{sec:integration}
Source: \texttt{integration.md} (draft)

\begin{itemize}
  \item No hard dependencies between tools --- every integration degrades gracefully
  \item L0 Presence: sentinel file check, no imports
  \item L1 Discovery: lazy, cached; missing binary = unavailable
  \item Owning tool is sole writer (never write across boundaries)
  \item Building blocks never check for consumers (one-way arrows)
\end{itemize}

% 
%   PART II: DEVELOPER WORKFLOW                                     
% 

\part{Developer Workflow}

% ──────────────────────────────────────────────────────────────────────
\section{Git \& Branching}
\label{sec:git}
Source: \texttt{workflow.md}

\begin{itemize}
  \item No direct commits to \texttt{main} --- all work on feature branches
  \item Prefixes: \texttt{feat/}, \texttt{fix/}, \texttt{refactor/}, \texttt{docs/}, \texttt{release/}, \texttt{chore/}
  \item One logical change per commit, 1--5 files, under 100 lines
  \item Conventional format: \texttt{type(scope): description}
  \item Quality gates pass before every commit (\texttt{make check})
\end{itemize}

% ──────────────────────────────────────────────────────────────────────
\section{Issue Tracking}
\label{sec:beads}
Source: \texttt{workflow.md}

\begin{itemize}
  \item \texttt{.beads/} initialized via \texttt{bd init}
  \item Claim before working: \texttt{bd update <id> --status in\_progress}
  \item Close when done: \texttt{bd close <id>}
  \item Work not complete until \texttt{git push} succeeds
  \item Session close: \texttt{git status}, commit, \texttt{bd sync}, push
\end{itemize}

% ──────────────────────────────────────────────────────────────────────
\section{Quality Gates \& CI}
\label{sec:ci}
Sources: \texttt{makefile.md}, \texttt{github.md}, \texttt{workflow.md}

\subsection{Makefile Targets (\must)}
\texttt{help}, \texttt{test}, \texttt{lint}, \texttt{check} (composes all),
\texttt{format}, \texttt{build}, \texttt{clean}. All in \texttt{.PHONY},
each with \texttt{\#\#} comment. \texttt{lint} never mutates.

\subsection{GitHub Branch Protection (\must)}
Require PR, conversation resolution, status checks. Prevent force push
and branch deletion. Zero bypass actors.

\subsection{CI Workflows (\must)}
Language-appropriate lint + test. \texttt{docs.yml} with markdownlint.
Action versions pinned to full SHA. Dependabot, secret scanning, push
protection enabled.

\subsection{Code Review (\must)}
PR via MCP $\to$ Copilot review $\to$ CI watch $\to$ address every comment
$\to$ 2--6 cycles $\to$ merge via MCP. Pre-PR local review with
\texttt{code-reviewer} + \texttt{silent-failure-hunter}.

% ──────────────────────────────────────────────────────────────────────
\section{CHANGELOG \& Docs}
\label{sec:changelog}
Source: \texttt{workflow.md}

\begin{itemize}
  \item Entries in PR branch before merge, under \texttt{[Unreleased]}
  \item Categories: Added, Changed, Fixed, Removed (omit empty)
  \item README updated when user-facing behavior changes
  \item Design decisions logged in \texttt{DESIGN.md} with reasoning and alternatives
\end{itemize}

% ──────────────────────────────────────────────────────────────────────
\section{Shell Scripts}
\label{sec:shell}
Source: \texttt{shell.md}

\begin{itemize}
  \item shellcheck: zero warnings, zero errors
  \item \texttt{\#!/usr/bin/env bash} + \texttt{set -euo pipefail}
  \item Double-quote all variable expansions and command substitutions
  \item Install scripts: \texttt{\#!/bin/sh} (POSIX), shellcheck with \texttt{--shell=sh}
\end{itemize}

% ──────────────────────────────────────────────────────────────────────
\section{Language Standards}
\label{sec:lang}

\subsection{Python}
Source: \texttt{python.md}.
Python 3.13+, uv, ruff, mypy strict, pyright strict, pytest, typer, FastMCP.
\texttt{from \_\_future\_\_ import annotations} in every file.
\texttt{internal/} for everything; core never imports from CLI.
No \texttt{Any} without documented reason.

\subsection{Go}
\texttt{internal/} for everything.
No \texttt{interface\{\}} or \texttt{any} unless unavoidable.
Errors are values: \texttt{fmt.Errorf("context: \%w", err)}.
Table-driven tests with testify. \texttt{-race} mandatory. Never log secrets.

\subsection{Node.js}
Source: \texttt{node.md}.
Node 20+, npm, ES modules, zod for validation,
\texttt{@modelcontextprotocol/sdk} for MCP.

% ──────────────────────────────────────────────────────────────────────
\section{Logging}
\label{sec:logging}
Source: \texttt{logging.md}

Logs for diagnosis, not monitoring. Never log content, messages, or API keys.
Location: \texttt{\textasciitilde/.punt-labs/<tool>/logs/}. MCP server logs to
stderr only (stdout reserved for stdio transport). Per-module logger via
\texttt{logging.getLogger(\_\_name\_\_)}. 5\,MB rotation, 5 backups.

% ──────────────────────────────────────────────────────────────────────
\section{Release Process}
\label{sec:release}
Sources: \texttt{release-process.md}, \texttt{release-requirements.md}

11 phases: Preflight $\to$ Version bump $\to$ Build $\to$ Release PR $\to$
Tag $\to$ CI wait $\to$ GitHub Release $\to$ PyPI verify $\to$
Post-release $\to$ Propagate $\to$ Verify.

\begin{itemize}
  \item All version locations match before tagging
  \item CHANGELOG stamped with date; new \texttt{[Unreleased]} inserted
  \item install.sh SHA updated in post-release
  \item Sibling repos updated (install-all.sh, marketplace, website)
  \item Idempotent: each step checks for diff before committing
\end{itemize}

\end{document}
